Outline

### 2021-10-05 text/main.Rmd
updates comments Vale, Jordan, Simon, Seb, Diego before resubmission

**Abstract:**
Preserving coral reef functioning is a critical challenge of the 21st century. However, a lack of quantitative assessments of multiple functions across large spatial scales has hindered local and regional conservation efforts. We integrate empirically-parametrized bioenergetic models and global community surveys to quantify five key functions mediated by coral reef fishes. We show that functions exhibit critical trade-offs driven by varying community structures, such that no community can maximize all functions. Further, functions are locally dominated by few species, but worldwide, the identity of dominant species greatly varies; more than half of the 1,110 species in our dataset are functionally dominant on a local scale. Our results underline the need for a nuanced, local, and context-dependent approach to coral reef conservation that considers multiple functions beyond the effect of standing stock biomass.

**Main Text:**
The flow of elements through biological communities fuels all ecosystems on earth [@Welti2017]. Harmful human dominance disrupts the natural functioning in contemperory ecosystems [@Ceballos2015].
Coral reefs represent a prime example of an ecosystem that is severely impacted by anthropogenic threats such as warming, and intensive fishing pressure [@Hughes2017; @Pauly2002].
Drastic declines in habitat quality and fish biomass have evoked serious concerns about the persistence of coral reef functioning [@Bellwood2019; @Williams2019; @Brandl2019front).
However, our knowledge on process-based functions - fluxes of elements and energy - is still in its infancy and evaluations of coral reef functioning typically rely on static proxies, such as live coral cover, standing stock biomass of reef fishes, or measures of diversity [e.g. @Cinner2020; @Mouillot2014; @Mora2011].
This sparsity of quantification of functions constitutes a severe limitation to efficient management of coral reef ecosystem functioning [@Brandl2019front].

Coral reef fishes play an essential role in coral reef functioning, because they contribute to a high proportion of consumer biomass and therefore represent essential vectors of fluxes of carbon (C), nitrogen (N) and phosphorus (P) through storage and recycling [@McIntyre2008; @Allgeier2014; @Barneche2014].
Ecosystem functions mediated by coral reef fishes include nutrient cycling, biomass production, herbivory , and piscivory (secondary consumption) (@Brandl2019front).
While the high diversity of coral reef fishes has inspired many studies with a focus on ecosystem functioning, only a handful of studies have attempted to quantify functions as continuous fluxes [@Brandl2019front]. Further, studies that do quantify functions mostly focused on biomass production (e.g., @Morais2019; @Morais2020glob) or fish excretion (e.g., @Allgeier2014), and suggest that functions rely on community structure. An assessment of multiple functions on a large scale is currently lacking.
Consequently, trade-offs amongs functions, their drivers, and their vulnerability to global stressors remain poorly understood in coral reef ecosystems [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the elemental fluxes that underpin reef fish functioning.
Using empirically-collected species-specific data on basic organismal processes and Bayesian phylogenetic models, we parameterize individual-level bioenergetic models to estimate five key ecosystem functions: N excretion, P excretion, biomass production, herbivory, and piscivory.
We apply these bioenergetic models to 9,118 reef fish transects across 585 sites worldwide (Table S1) to:
(1) quantify community-level reef fish functions and their trade-offs,
(2) extract the community- and species-level effects on these functions, and
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

## Results
By combining individual-level bioenergetic models and global underwater visual transect data, we quantified five key ecosystem functions mediated by coral reef fishes across the globe. We estimated the average functions per locality as well as a combined measure of multifunctionality by running a multivariate Bayesian regression model (Fig. 1). Functions across localities show similar spatial patterns, yet a high level of function does not nescessarily ensure a high level of another function. For example, Trindade in the South Atlantic Ocean has high values for N excretion, P excretion, and piscivory, but average values for biomass production. A measure of multifunctionality thus does not appropriately represent the state of all functions.

Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we demonstrate a predictably strong relationship between fish biomass and all five functions by running multivariate mixed effect Bayesian model (Fig. 1). Specifically, the effects of the log-transformed biomass on log-transformed functions were 0.932 (95\%CI: 0.929; 0.934), 1.051 (1.047; 1.056), 0.771 (0.764; 0.780), 0.940 (0.923; 0.957), and 0.668 (0.635; 0.702) for N excretion, P excretion, production, herbivory, and piscivory, respectively. We also incorporated sea surface temperature (sst) because it affects the metabolic and growth rates of fishes, which scales up to the community [@Barneche2019]. We found positive effects of sst on N excretion, production, and herbivory, and negligible effects of sst on P excretion and piscivory (table Sx).

The above-mentioned multivariate model also allowed us to estimate the correlations between functions, independent of the effects of biomass and sst.
Specifically, we estimated correlations between functions on three levels: the locality effects, the site effects, and the residual variations (Fig. 2, fig. Sx.
The correlations mostly overlapped across the three levels for a given pair of functions. We found consistently negative trade-offs between P excretion and N excretion, P excretion and biomass production. Further, there were slightly weaker negative correlations between piscivory and N excretion and between piscivory and herbivory. As a consequence of the negative correlations, a fish community cannot have high values for all functions simultaneously.

To explore what drives the residual variation and trade-offs of functions beyond the effects of biomass and sst, we conducted a multivariate Bayesian mixed effect model by adding ten variables that describe the community structure: species richness and the median, lower, and upper 95% quantiles of size, immaturity, and trophic level of the individuals inside a community. We found that the community structure drives the observed trade-offs (Fig. 3; table x).

We then questioned if functions tend to be driven by specific species across sites. We quantified the degree of dominance inside each community on the site-level and found that functions are on average charachterized by a high level of dominance, i.e. in most communities, a small proportion of species perform most of a function (Fig. 4a). As a comparison, we simulated random communities ranging from 5 to 50 species which yielded an average XX. The predicted average of the degree of dominance was x, x, x, x, and x for each function respectively.
We also calculated the proportion of species that is dominant in at least one transect (i.e. species that contribute more than a community in which each species contributes equally). We found that most species play a dominant role, and 70\% of all species contributed disproportionally to a function in at least one community (Fig. 4b).
Finally, we investigated how often species are dominant across their range and found that very few species are dominant throughout their range (Fig. 4c). Together, this means that functions inside communities tend to be driven by few dominant species, but that the identity of these dominant species varies highly across communities.

@@

By quantifying five key functions mediated by coral reef fishes, we demonstrate that ecosystem-wide functioning is mainly driven by fish biomass, yet biomass alone does not fully capture functional variability across communities.
While biomass is one of the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], our analyses suggests that biomass alone does not sufficiently characterize functioning.
We demonstrate strong trade-offs among the five functions, independent of biomass. Using biomass as a proxy for functioning can therefore mask fundamental differences in critical community-level functions.
Furthermore, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize all functions simultaneously.

Fish community structures drive the observed trade-offs among functions which can be explained by organismal physiology and life-history traits of the individuals inside communities [@Schiettekatte2020; @Schramski2015].
For example, we observed a clear trade-off between P excretion and biomass production which is largely driven by the age and trophic community structure.
Communities dominated by fishes with high trophic levels are charachterized high P excretion rates because predatory fishes have a P-rich diet (@Allgeier2014).
In contrast, biomass production is high in communities dominated by fishes with low trophic levels because herbivores tend to have higher growth rates [@Morais2018].
Moreover, communities dominated by mature fishes have high values of P excretion because growing fishes are more likely to retain P to form P-rich bones [@Schiettekatte2020; @Hood2005], while communities dominated by immature fishes maximize biomass production.
Furthermore, metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportionate contributions to functions that rely on rapid energetic turnover such as herbivory, piscivory, production, and N excretion [@Barneche2018; @Morais2020glob; @Brandl2019science].

Our results also reveal that functions consistently hinge on a few dominant species, but the identities of local, dominant species vary strongly across sites [@Lefcheck2019].
Locally, specific high-performing taxa may disproportionately impact rates of functioning at the community level due to high biomass or abundance [@Topor2019].
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
While few high-performing taxa indeed dominate functioning in each site, there are no species that are dominant across their entire range.
In addition, more than half of all species contributed disproportionally to a specific function in at least one site inside their range.
There are thus no widespread key species to target for conservation, rather, the functional role and conservation priority of species should be investigated at a local scale.

Our global analysis of multiple functions suggests pathways in which human-induced shifts in reef fish community structure may impact coral reef ecosystems (Fig. 5).
Fishing and climate-induced coral loss have caused declines in reef fish biomass and shifts in community structure [@Darling2017; @Graham2017], and our findings suggest that these changes will have effects on ecosystem functioning.
Intensive fishing and associated reductions in biomass of large fishes alters the size, age, and trophic structure of fish communities [@Graham2017].
When accounting for the effect of biomass, these community shifts can enhance nitrogen excretion and production [e.g. @Morais2020glob], but negatively impact phosphorus excretion, herbivory, and piscivory.
Futher, declines in coral cover related to climate change are often associated with a shift toward herbivores, which may deter algal domination [@Graham2006].
Simultaneously, herbivores have a minor contribution to P excretion [@Allgeier2014; @Schiettekatte2020], so a shift to herbivore dominance and the subsequent decline of community-level P excretion may change the balance of nutrient cycling on coral reefs, potentially favoring algal growth in comparison with corals [@Burkepile2013].

Conserving biomass, diversity, and ecosystem functions are important objectives of most conservation initiatives [@Cinner2020].
While safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory.
However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards herbivore domination may result in the decline of P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions, yet our comprehension of
process-based functioning of coral reefs or what a “functional” coral reef really means is still in its infancy [@Brandl2019front].
To appropriately manage coral reef functioning with respect to local needs and stressors, future studies should seek to establish functional baselines for global coral reefs.
Until then, our study suggests that coral reef fish functions can be managed by enhancing standing stock biomass, protecting local key species, and promoting a balanced community structure.

Overall, using coral reef fishes, we demonstrate that the variability in processes that govern the elemental cycling presents an unrecognized challenge for protecting ecosystem functioning. Management strategies that call for the enhancement of coral reef functioning via an economic mindset (i.e. where higher functioning is better) are not feasible. Instead, conserving coral reef ecosystem functioning will require a more nuanced approach that considers processes that vary beyond the effect of standing stock biomass and are subject to variable, local trade-offs, drivers, and anthropogenic threats.

@@

**Acknowledgements:**
We thank the staff at CRIOBE, Moorea for field support. We would also like to thank Jérémy Carlot, Beverly French, Titouan Roncin, Yann Lacube, Camille Gache, Gabrielle Martineau, Kailey Bissell, Benoit Espiau, Calvin Quigley, Kaitlyn Landfield and Tommy Norin for their help in the field, Sophie Schiettekatte for proof-reading the manuscript, and Guillemette de Sinéty and Jérémy Wicquart for their contribution to otolith analyis.
This research was funded by the BNP Paribas Foundation (Reef Services Project) and the French National Agency for Scientific Research (ANR; REEFLUX Project; ANR‐17‐CE32‐0006). This research is product of the SCORE-REEF group funded by the Centre de Synthèse et d’Analyse sur la Biodiversité (CESAB) of the Foundation pour la Recherche sur la Biodiversité (FRB) and the Agence Nationale de la Biodiversité (AFB). VP was supported by the Institut Universitaire de France (IUF) and JMC was supported by a Make Our Planet Great Again Postdoctoral Grant (mopga‐pdf‐0000000144).
**Author contributions:**
NMDS and VP conceived the idea and NMDS, VP, SJB, and JMC designed methodology; NMDS, JMC, SJB, AM, FM, VP, KSM, JEA and DEB collected the data; All authors shared existing data. NMDS analyzed the data and led the writing of the manuscript. All authors contributed significantly to the drafts and approved the final version for publication.
**Competing interests:** None declared.
**Data and materials availability:**
All data and code to reproduce figures are available online through GitHub (https://github.com/nschiett/global_proc) and figshare (https://figshare.com/s/f789aec2c20492c4f0f9). All data on individual fish traits are available from the corresponding author on reasonable request.

@@

Fig. 1: Maps of five key ecosystem functions and multifunction, and their relationships with biomass. Left: Dots indicate localities of field surveys, with dot sizes representing the ranked values of the locality-level predictions of functions, and color scales showing categorical assignments (black = lower 25\%, grey = 25-75\%, color = >75\%). Black circular outlines highlight the five localities with the highest values of each function. Multifunction represents the weighted average of the five standardized functions. Right: The predicted values for functions and multifunction with increasing biomass. The lines represent the average predicted value and the shaded areas show the 95% credible interval of he predictions.

Fig. 2: Correlations among five functions while accounting for biomass and sea surface temperature. a: Modeled correlation coefficients of residual errors. Dots represent the average and lines represent the 95%CI. b-k: Scatter plots of the mean residual error of the functions to demonstrate the correlations.

Fig. 3.  Effects of ecological community variables on five functions. Dots indicate fixed effect values from Bayesian linear regressions that examine effects of species richness, trophic level, size, and immaturity of fishes. To represent both the median and the spread of trophic level, size, and immaturity across individuals inside a community, we included lower and upper 95% quantile values of these three traits as community variables. All data were log-transformed and standardized to compare across functions and variables (see Table S2 for parameter values on non-standardized data). Dots represent the average effect size estimate, and horizontal lines indicate the 95% credible interval. Immaturity is defined as the derivative of the Von Bertalanffy growth model for a given size, and the higher this value, the younger the individual.

Fig. 4: Local dominance in species contributions for five ecosystem functions on coral reefs. a: The degree of dominance of communities for each function on the site-level. Colored dots represent the raw values and the black dots and lines display the mean and 95\% credible intervals of the average degree of dominance, respectively. Degrees of dominance range between zero (all species contribute equally) and one (a single species is the sole contributor to a given function). The vertical dashed line shows the average of the degree of dominance of 1,000 randomly simulated communities. b: Bar plot showing the proportion of species that are dominant in at least one site relative to the total number of species. In the case of herbivory and piscivory, the total number of herbivores and piscivores, respectively, is used.  c: Species-specific frequencies of dominance in each function across all sites, ranging from zero (species are never dominant) to one (dominant wherever present). A species is categorized as dominant in a community if its contribution to a function is higher than a scenario in which all species are equal (i.e. one divided by the number of species that contribute to the function).

Fig. 5. Vulnerability of five critical functions to fishing and climate change-induced coral loss. Conceptual scheme showing the potential ways in which fishing and climate change may affect functions through their effects on biomass and community structure.

>>>>

**Abstract:**
Preserving coral reef functioning is a critical challenge of the 21st century. However, a lack of quantitative assessments of multiple functions across large spatial scales has hindered local and regional conservation efforts. We integrate empirically-parametrized bioenergetic models and global community surveys to quantify five key functions mediated by coral reef fishes. We show that functions exhibit critical trade-offs driven by varying community structures such that no community can maximize all functions. Further, functions are locally dominated by few species, but the identity of dominant species substantially varies at the global scale. In fact, more than half of the 1,110 species in our dataset are functionally dominant on a local scale. Our results reinforce the need for a nuanced, locally tailored approach to coral reef conservation that considers multiple ecological functions beyond the effect of standing stock biomass.

**Main Text:**
The flow of elements through biological communities fuels all ecosystems on earth [@Welti2017]. Humans increasingly threaten biodiversity and ecosystem functioning [@Ceballos2015].
Coral reefs are a prime example of an ecosystem severely impacted by anthropogenic activities.
Drastic declines in habitat quality and fish biomass have evoked serious concerns about the persistence of coral reefs [@Hughes2017; @Pauly2002].
Maintaining ecosystem functions, defined as fluxes of elements, is a major goal for conservation of coral reefs [@Bellwood2019; @Williams2019; @Brandl2019front].
However, past evaluations of functions on coral reefs have mostly relied on static proxies such as live coral cover, standing stock biomass of reef fishes, or measures of diversity [@Cinner2020; @Mouillot2014; @Mora2011].
These simplified proxies, although useful, may not properly represent ecological functions because the fluxes of elements can scale non-linearly with variables such as biomass [@Barneche2014].
Therefore, improving the quantification of ecological functions constitutes an important step towards the efficient management of coral reef ecosystem functioning [@Brandl2019front].

As a dominant group of consumers, coral reef fishes are essential vectors of carbon (C), nitrogen (N) and phosphorus (P) [@McIntyre2008; @Allgeier2014; @Barneche2014].
Ecosystem functions mediated by coral reef fishes include nutrient cycling, biomass production, herbivory, and piscivory (secondary consumption) (@Brandl2019front).
While the high diversity of coral reef fishes has inspired many studies that focus on ecosystem functioning, only a handful of studies have attempted to actually quantify functions as continuous fluxes [@Brandl2019front]. Further, studies that quantify functions as a flow of matter mostly focus on single functions (e.g. biomass production [@Morais2019; @Morais2020glob] or fish excretion [ @Allgeier2014]) and are restricted to a small number of species and local scales.
Consequently, trade-offs amongs multiple functions, their drivers, and their vulnerability to anthropogenic stressors remain poorly understood in coral reef ecosystems across large spatial scales [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the elemental fluxes that underpin reef fish functioning.
Using empirical species-specific data on basic organismal processes and Bayesian phylogenetic models, we parameterize individual-level bioenergetic models to estimate five key ecosystem functions: N excretion, P excretion, biomass production, herbivory, and piscivory for 1,100 species.
We apply these bioenergetic models to 9,118 reef fish transects across 585 sites worldwide (Table S1) to:
(1) quantify community-level reef fish functions and their trade-offs,
(2) extract the community- and species-level effects on these functions, and
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

## Results
We quantified five key ecosystem functions mediated by coral reef fishes across the globe (Fig. 1). Functions across localities show a similar geographocal pattern. However, at the global scale we could not find a single location with high levels of functioning across all functions. A measure of multifunctionality thus does not appropriately represent the state of all functions.

Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we demonstrate a predictably strong relationship between fish biomass and all five functions (Fig. 1). Specifically, in a multivariate mixed effects Bayesian model, the slopes of the log-transformed biomass were 0.932 (95\%CI: 0.929, 0.934) for N excretion, 1.051 (1.047, 1.056) for P excretion, 0.771 (0.764, 0.780) for production, 0.940 (0.923, 0.957) for herbivory, and 0.668 (0.635, 0.702) for piscivory. These slopes indicate that the relationships between biomass and functions are all non-linear, which demonstrates that biomass is not an appropriate proxy for function. We also incorporated sea surface temperature (SST) because it affects the metabolic and growth rates of fishes, which scales up to the community [@Barneche2019]. We found positive effects of SST on N excretion, production, and herbivory, and no effects of SST on P excretion and piscivory (table S3).

The forementioned multivariate model also allowed us to estimate the correlations between functions, independent of the effects of biomass and SST.
In particular, we estimated correlations between functions on three levels: the locality effects, site effects, and residual variations (Fig. 2, Fig. S1).
The correlations displayed comparable patterns on each level.
We found strong negative trade-offs between P excretion and N excretion as well as P excretion and biomass production. Further, we found slightly weaker negative correlations between piscivory and N excretion as well as piscivory and herbivory. Thus, a reef fish community does not simultaneously display high values of functioning across all investigated functions.

To explore what drives the variation and trade-offs of functions beyond the effects of biomass and SST, we ran a multivariate Bayesian mixed effects model by adding ten variables that describe the structure of each fish assemblage: species richness and the median, lower, and upper 95% quantiles of size, immaturity, and trophic level of individuals inside a community.
Each of these components have non-zero effects on at least one of the five functions, suggesting that the observed trade-offs may be, at least in part, rooted in the structure of the focal community (Fig. 3; Table S4).

Beyond community structure, we examined whether functions are driven by specific species across sites. We quantified the degree of dominance inside each community at the site-level and found that, on average, functions are dominated by a small fraction of the species in each community (Fig. 4a).
We also calculated the proportion of species that is dominant in at least one transect (i.e. species with a disproportionally high contribution as compared to a community in which all species contribute equally). However, we also found that at a global scale many species play a dominant role. In particular, 70\% of all species contributed disproportionally to a function in at least one community (Fig. 4b).
Yet, very few species are dominant throughout their range (Fig. 4c). Thus, functions within communities tend to be driven by few dominant species, but the identity of those dominant species varies across sites.

@@

By quantifying five key processes mediated by coral reef fishes, we demonstrate that coral reef ecosystem functioning is shaped by biological trade-offs, local community structure, and species identity. Standing biomass is one of the most commonly employed indicators of coral reef functioning [@Cinner2020; @Brandl2019front], and our analyses confirm the pervasive influence of biomass on all other processes. Yet, our results also show non-linear relationships between functions and biomass and suggest that biomass alone does not sufficiently characterize functioning, as we found strong trade-offs among the five functions independent of biomass. Using biomass as a proxy might, therefore, mask fundamental differences in community-level functioning. Furthermore, for a given value of biomass, no reef can yield above average values across all five functions. While a reef may stand out as a hotspot for one function, no reef can maximize all functions simultaneously.

The observed trade-offs among functions are driven by fish community structure and the organismal physiology and life-history traits of its consituents [@Schiettekatte2020; @Schramski2015].
For example, we observed a clear trade-off between P excretion and biomass production which is mostly driven by the age and trophic structure.
Communities dominated by fishes with high trophic levels are charachterized high P excretion rates because predatory fishes have a P-rich diet [@Allgeier2014].
In contrast, biomass production is high in communities dominated by fishes that occupy low trophic levels because herbivores tend to exhibit higher growth rates [@Morais2018].
Moreover, P is retained for skeletal growth in young fishes, thus limiting P excretion rates [@Schiettekatte2020; @Hood2005].
Furthermore, metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportionate contributions to functions that rely on rapid energetic turnover such as herbivory, piscivory, production, and N excretion [@Barneche2018; @Morais2020glob; @Brandl2019science].

Our results also reveal that functions consistently rely on a few dominant species, but the identities of local, dominant species strongly vary across sites [@Lefcheck2019].
Locally, a small number of high-performing taxa may disproportionately impact rates of functioning at the community level due to high biomass or abundance [@Topor2019], which may have led to their designation as functionally-dominant “key species” in various locations [@Bellwood2006].
However, our results revealed that no species dominated throughout their geographical range, and more than half of all species contributed disproportionally to a specific function at a at least one site.
Thus, there are no widespread key species to target for coral reef conservation; rather, local species dominance across functions can guide local conservation efforts, and the preservation of regional reef fish biodiversity should be prioritized in broad-scale policy.

Our global analysis of multiple functions suggests pathways in which human-induced shifts in reef fish community structure may impact coral reef ecosystems (Fig. 5).
Fishing and climate-induced coral loss have caused declines in reef fish biomass and shifts in community structure [@Darling2017; @Graham2017], and we suggest that these changes will differentially affect ecosystem functioning.
Intensive fishing and associated reductions in the biomass of large fishes, for example, alters the size, age, and trophic structure of fish communities [@Graham2017].
When accounting for the effect of biomass, these community shifts can enhance N excretion and production [e.g. @Morais2020glob], but they will negatively impact P excretion, herbivory, and piscivory.
Further, declines in coral cover related to climate change are often associated with shifts toward herbivores [@Graham2006].
Herbivores generally contribute little to P excretion [@Allgeier2014; @Schiettekatte2020], so a shift to herbivore dominance and the subsequent decline of community-level P excretion may change the balance of nutrient cycling on coral reefs, potentially favoring algal growth over corals [@Burkepile2013].

Sustaining biomass, diversity, and ecosystem functioning are important objectives of most conservation initiatives [@Cinner2020].
While safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation stratigies for coral reefs [@Graham2020], may protect herbivorous species.
However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards herbivore domination may result in the decline of P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions.
Yet, our comprehension of process-based functioning or the definition a “functional” coral reef is still poorly understood [@Brandl2019front].
Establishing functional baselines for global coral reefs is a critical challenge for future studies.
Until then, our results suggest that coral reef fish functions can be managed by enhancing standing stock biomass, protecting local key species and vulnerable consituents of the community (e.g. large carnivores), and promoting regional biodiversity.

We demonstrate that the variability in processes that govern the elemental cycling in complex ecosystems such as tropical coral reefs represents an unrecognized challenge for protecting ecosystem functioning. Management strategies that call for the enhancement of ecosystem functioning via an economic mindset (i.e. where higher functioning is better) are not feasible. Instead, conserving coral reef ecosystem functioning will require a more nuanced approach that considers processes that vary beyond the effect of standing stock biomass and are subject to local trade-offs, drivers, and anthropogenic threats.

@@

**Acknowledgements:**
We thank the staff at CRIOBE, Moorea for field support. We would also like to thank Jérémy Carlot, Beverly French, Titouan Roncin, Yann Lacube, Camille Gache, Gabrielle Martineau, Kailey Bissell, Benoit Espiau, Calvin Quigley, Kaitlyn Landfield and Tommy Norin for their help in the field, Guillemette de Sinéty and Jérémy Wicquart for their contribution to otolith analysis, and Sophie Schiettekatte for proof-reading the manuscript. .
This research was funded by the BNP Paribas Foundation (Reef Services Project) and the French National Agency for Scientific Research (ANR; REEFLUX Project; ANR‐17‐CE32‐0006). This research is product of the SCORE-REEF group funded by the Centre de Synthèse et d’Analyse sur la Biodiversité (CESAB) of the Foundation pour la Recherche sur la Biodiversité (FRB) and the Agence Nationale de la Biodiversité (AFB). VP was supported by the Institut Universitaire de France (IUF) and JMC was supported by a Make Our Planet Great Again Postdoctoral Grant (mopga‐pdf‐0000000144).
**Author contributions:**
NMDS and VP conceived the idea and NMDS, VP, SJB, and JMC designed methodology; NMDS, JMC, SJB, AM, FM, VP, KSM, JEA and DEB collected the data; All authors shared existing data. NMDS analyzed the data and led the writing of the manuscript. All authors contributed significantly to the drafts and approved the final version for publication.
**Competing interests:** None declared.
**Data and materials availability:**
All data and code to reproduce the figures are available on GitHub (https://github.com/nschiett/global_proc) and figshare (https://figshare.com/s/f789aec2c20492c4f0f9). All data on individual empirical measurements are available from the corresponding author upon request.

@@

Fig. 1: Maps of the five key ecosystem functions, multifunctionality, and the relationships between the functions and biomass. Left: Dots indicate localities of field surveys, with dot sizes representing the ranked values of the locality-level predictions of functions, and color scales showing categorical assignments (black = < 25\%, grey = 25-75\%, color = >75\%). Black outlines highlight the five localities with the highest values of each function. Multifunctionality represents the weighted average of the five standardized functions. Right: The predicted values for functions and multifunctionality with increasing biomass. The lines represent the average modeled relationship and the shaded areas show the 95\% credible intervals of the predictions.

Fig. 2: Correlations of the five functions, accounting for biomass and sea surface temperature. a: Modeled correlation coefficients of residual errors. Dots represent the average and lines represent the 95%CI. b-k: Scatter plots of the mean residual errors of the functions.

Fig. 3.  Effects of ecological community variables on the five functions. Dots indicate fixed effect values from Bayesian linear regressions that examine the effects of species richness, trophic level, size, and immaturity of fishes. To represent both the median and spread of trophic level, size, and immaturity across individuals within a community, we included lower and upper 95% quantile values of these three traits as community variables. All data were log-transformed and standardized to compare across functions and variables (see Table S2 for parameter values on non-standardized data). Dots represent the average effect size estimate, and horizontal lines indicate the 95% credible interval. Immaturity is defined as the derivative of the von Bertalanffy growth model for a given size; thus, the higher this value, the younger the individual.

Fig. 4: Local dominance in species contributions to five ecosystem functions on coral reefs. a: The degree of dominance for each function at the site level. The degree of dominance of a community ranges between zero (all species contribute equally to the function) and one (a single species is the sole contributor to a given function). Colored dots represent the raw values, and the black dots and lines display the mean and 95\% credible intervals of degree of dominance among all sites. In some cases, the credible interval was too small to be visible. The vertical dashed line shows the average degree of dominance of 1,000 randomly simulated communities. b: Bar plot of the proportion of species that are dominant in at least one site relative to the total number of species, or, for herbivory and piscivory, the total number of herbivores and piscivores, respectively. c: Species-specific frequencies of dominance in each function across all sites, ranging from zero (species are never dominant) to one (dominant wherever present). A species is categorized as dominant in a community if its contribution to a function is higher than a scenario in which all species are equal (i.e. one divided by the number of species that contribute to the function).

Fig. 5. Vulnerability of the five functions to fishing and climate change-induced coral loss. Conceptual scheme of the potential ways in which fishing and climate change can affect functions through their known effects on biomass and community structure.
